---
title: "Initial temperature trend exploration: Yolo County"
output: github_document
---
```{r setup,echo=FALSE}
library(knitr)
knitr::opts_chunk$set(echo=FALSE)
setwd("~/project_repo/birds_precip/")
```
This data was obtained from the NOAA Climate Data Online archive (https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/locations/FIPS:06113/detail). There are 19 stations in the Yolo County:

```{r read_info}
yolo_data=read.csv("yolo_county_2000_2019.csv")
tab_info=unique(yolo_data[c("NAME","LATITUDE","LONGITUDE")])
kable(tab_info,row.names=FALSE)
```

For now, we are looking at data from January 2000-June 2019, although we can grab more data in the future if needed. Note that not all stations provide the same amount of data!

Listed fields in the CSV:

- TOBS: Temperature at observation time
- TMAX: Max temperature
- TMIN: Min temperature
- TAVG: Average temperature
- PRCP: Precipitation
- MDPR: Multi-day precipitation total
- DAPR: Number of days in MDPR
- SNOW: snowfall
- SNWD: Snow depth
- WT01: Fog
- WT03: Thunder 
- WT05: Hail
- WT11: High or damaging winds
- MDWM: Multiday wind movement
- DAWM: Number of days in MDWM
- WDMV: Total wind movement
```{r available_data}
stations=unique(yolo_data$NAME)
for (s in stations){
  yolo_subset=yolo_data[yolo_data$NAME==s,]  
  #Which columns are non-NA?
  has_values<-c()
  for (c in names(yolo_subset)[7:22]){
    if (all(is.na(yolo_subset[,c]))==FALSE){
      has_values<-c(has_values,c)
    }
  }
  print(sprintf("Station %s has %s",s,paste(has_values,collapse = ", ")))
}



```